---
title: "Our Mindfulness Community"
author: "Carey Huh, Ph.D."
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(readxl)
library(lubridate)
library(viridis)

source('theme_carey.R')
```

## Project Background

* Susan Samueli Integrative Health Institute has graciously offered the UCI and the community opportunities to practice mindful meditatations for two years now

* My name is Carey Huh and I worked at UCI as a neuroscientist for over 5 years

* My mindfulness journey began with a free MBSR course offered by SSIHI in May 2020, at the start of the COVID-19 pandemic

* Since then, I have been a regular practitioner at SSIHI's weekday community meditations and yoga for wellbeing classes 

* In April 2021, I participated in the 8-week Mindful Self-Compassion course offered by SSIHI, taught by Amy Noelle

* The consistent mindfulness practices and self-compassion teachings helped to completely transform my outlook on life!

* I changed my career to a data scientist working for the CDC Foundation

* A career that allows me to work from home means that I can stay connected to the UCI mindfulness community that I have grown to love and cherish

## Project Motivation

* As a passion project, I have been monitoring attendance at SSIHI mindfulness community classes since April of 2021

* This attendance data only contain aggregated summary statistics and the anonymity of participants is strictly protected

* Using the data I have collected, I wanted to take the opportunity to show how the community has stayed consistent in number throughout the length of the program

* In particular, individual engagement during community meditation classes has steadily increased, indicating that individual participation and willingness to share in the sense of community has only grown stronger, not weaker, over time

```{r import_data}
df <- read_excel("mindful_database.xlsx") %>%
  mutate(Time_Start_24hr = substr(Time_Start_24hr, 12, 16),
         Date_Monday = floor_date(Date, "weeks", week_start = 1)) %>%
  select(Number, Date, Date_Monday, Day_Of_Week, everything())

free_meditation <- df %>%
  filter((Practice_Type == "Meditation") & (Cost == "Free"))

yoga <- df %>%
  filter(Practice_Type == "Yoga")

paid_meditation <- df %>%
  filter((Practice_Type == "Meditation") & (Cost != "Free"))

other <- df %>%
  filter(!(Practice_Type %in% c("Meditation","Yoga")))
```


## Weekday Community Meditation

```{r meditation_attend}
free_meditation %>%
  group_by(Date_Monday) %>%
  summarise(attend = as.integer(mean(Participants_Include_Instructor))) %>%
  ggplot(aes(x = Date_Monday, y = attend)) +
  geom_col(aes(fill = as.factor(quarter(Date_Monday))), color = "black") +
  scale_fill_viridis(discrete = TRUE) +
  theme_gray() +
  labs(
    title = "Attendance at Community Meditation Has Been Consistent Over Time",
    tag = "Figure 1",
    x = "",
    y = "Number of Participants Per Class",
    caption = "Each bar represents a weekly average statistic",
    fill = "Quarter")
```



