---
title: "Mindfulness Meditation Transforms Lives"
subtitle: "Evidence-Based Findings & Testimonials from Our Mindfulness Communiity"
author: "Carey Huh, Ph.D."
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

library(tidyverse)
library(readxl)
library(lubridate)
library(viridis)
library(googlesheets4)

source('theme_carey.R')

google_url <- "https://docs.google.com/spreadsheets/d/1RLOJXi2VQBOabrcT1qiLiyNO1k7eW8wm7E1Fze1NlRM/edit?usp=sharing"
```

## Background

* Susan Samueli Integrative Health Institute has graciously offered UCI and the greater community daily opportunities to practice mindfulness for roughly two years

* Consistent mindfulness practices have improved mental and physical well-being of community members and have transformed lives (please see **Testimonials** section below)

* As a data scientist, I have been curious about how the mindfulness community has evolved over time 
 
* To gather data for this passion project, I have been monitoring attendance at SSIHI mindfulness community classes since April of 2021

* This attendance data only contain aggregated summary statistics and the anonymity of participants is strictly protected

* Using the data I have collected, I wanted to take the opportunity to show how the community has stayed consistent in number throughout the length of the program

* Moreover, individual engagement during community meditation classes has always been relatively high despite virtual format

* Data indicate that the mindfulness community is evolving; the percentage of male participants is steadily increasing

* Small class sizes encourage active participation by community members

```{r import_data}
df <- read_sheet(google_url) %>%
  mutate(Time_Start_24hr = substr(Time_Start_24hr, 12, 16),
         Date_Monday = floor_date(Date, "weeks", week_start = 1),
         Sharing_Participants = as.numeric(Sharing_Participants, na.omit = TRUE),
         Day_Of_Week = as.factor(Day_Of_Week)) %>%
  select(Number, Date, Date_Monday, Day_Of_Week, everything())
df$Day_Of_Week = fct_relevel(df$Day_Of_Week, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Sunday"))

free_meditation <- df %>%
  filter((Practice_Type == "Meditation") & (Cost == "Free"))

yoga <- df %>%
  filter(Practice_Type == "Yoga")

paid_meditation <- df %>%
  filter((Practice_Type == "Meditation") & (Cost != "Free"))

other <- df %>%
  filter(!(Practice_Type %in% c("Meditation","Yoga")))
```


## Weekday Community Meditation

```{r meditation_attend}
free_meditation %>%
  group_by(Date_Monday) %>%
  summarise(attend = as.integer(mean(Participants_Include_Instructor))) %>%
  # plot attendance
  ggplot(aes(x = as.Date(Date_Monday), y = attend)) +
  geom_col(aes(fill = as.factor(quarter(Date_Monday))), color = "black") +
  scale_fill_viridis(discrete = TRUE, direction = -1) +
  scale_x_date(date_breaks = "3 months", date_labels = "%Y-%b") +
  labs(
    title = "Attendance at Community Meditation Has Been Consistent Over Time",
    tag = "Figure 1",
    x = "Date (Year-Month)",
    y = "Number of Participants Per Class",
    caption = "Each bar represents a weekly average based on classes attended by the author.",
    fill = "Quarter") +
  theme_gray()
```

* Figure 1 shows that the number of participants at community mediation classes has been consistent over time

* The spike in attendance during the summer of 2021 (2021-Q3) is most likely due to the employee well-being program that temporarily allowed participation from employees of other UC campuses

```{r meditation_camera}
free_meditation %>%
  mutate(On = Camera_On_Participants / Participants_Include_Instructor * 100,
         Off = 100 - On) %>%
  mutate(Year_Quarter = paste(year(Date), quarter(Date), sep = "-Q")) %>%
  group_by(Year_Quarter) %>%
  summarise(On = mean(On),
            Off = 100 - On) %>%
  pivot_longer(cols = c(On, Off)) %>%
  # plot camera_on, camera_off
  ggplot(aes(x = Year_Quarter, y = value)) +
  geom_col(aes(fill = name), color = "black") +
  scale_fill_viridis(discrete = TRUE) +
  labs(
    title = "Percentage Participation with Camera On Remains High",
    tag = "Figure 2",
    x = "Date (Year-Quarter)",
    y = "Percentage of Participants",
    caption = "Each bar represents a quarterly average based on classes attended by the author.",
    fill = "Camera Status") +
  theme_gray()
```

* Whether a participant has their camera on during a virtual mediation class may indicate greater engagement, comfort and willingness to share 

```{r meditation_camera_on_percent}
free_meditation %>%
  mutate(On = Camera_On_Participants / Participants_Include_Instructor * 100) -> df_temp
value <- as.integer(mean(df_temp$On), NA.omit = TRUE) 
```


* Figure 2 shows that the percentage of participation with camera on has remained relatively stable across time

* The overall percentage of participation with camera on at community meditation classes is relatively high at `r value`%

```{r meditation_gender}
free_meditation %>%
  mutate(female = Female_Participants / Participants_Include_Instructor * 100,
         male = Male_Participants / Participants_Include_Instructor * 100,
         unknown = 100 - female - male) %>%
  mutate(Year_Quarter = paste(year(Date), quarter(Date), sep = "-Q")) %>%
  group_by(Year_Quarter) %>%
  summarise(Female = mean(female),
            Male = mean(male),
            Unknown = 100 - Female - Male, 
            Male_SEM = sd(male) / sqrt(n())) %>%
  pivot_longer(cols = c(Female, Male, Unknown)) %>%
  mutate(name = as.factor(name)) -> df_temp

df_temp$name <- fct_relevel(df_temp$name, levels = c("Unknown", "Female", "Male")) 

df_temp %>%
  filter(name == 'Male') -> df_temp1
  
df_temp %>%
  # plot gender
  ggplot(aes(x = Year_Quarter, y = value)) +
  geom_col(aes(fill = name), color = "black") +
  geom_errorbar(data = df_temp1, width = 0.5, 
                aes(x = Year_Quarter, ymin = value - Male_SEM, ymax = value + Male_SEM)) +
  scale_fill_viridis(discrete = TRUE) +
  labs(
    title = "Male Participation at Community Meditation Has Steadily Increased",
    tag = "Figure 3",
    x = "Date (Year-Quarter)",
    y = "Percentage of Participants",
    caption = "Each bar represents a quarterly average based on classes attended by the author. \n 
    Error bars represent mean ± standard error.",
    fill = "Gender") +
  theme_gray()
```

* It is a well-known trend that community meditation classes attract more women than men

```{r meditation_male_percent}
df_temp %>%
  filter((Year_Quarter %in% c('2021-Q2', '2022-Q4')) & (name == 'Male')) %>%
  pull(value) %>%
  as.integer() -> values
```

* Figure 3 suggests that the percentage of male participants has been steadily increasing over time

* The average percentage of male participants increased from `r values[1]`% to `r values[2]`% in 1.5 years

```{r meditation_day_of_week}
free_meditation %>%
  # plot day of the week attendance
  ggplot(aes(x = Day_Of_Week, y = Participants_Include_Instructor)) +
  geom_boxplot(outlier.shape = NA, show.legend = FALSE) +
  geom_jitter(aes(color = Day_Of_Week), width = 0.1, alpha = 0.8, show.legend = FALSE) +
  scale_color_viridis(discrete = TRUE) +
  theme(legend.position = "none") +
  labs(
    title = "Attendance at Community Meditation is Highest at Start of the Week",
    tag = "Figure 4",
    x = "Day of Week",
    y = "Number of Participants Per Class",
    caption = "Each jittered dot represents one class. Center line in boxplots represents the median.") +
  theme_gray()
```

* Figure 4 indicates that community meditation attendance is highest on Mondays and lowest on Wednesdays

* Days of the week with the highest attendance numbers are Mondays, Tuesdays and Fridays

* Note that meditation classes take place at noon on Mondays, Wednesdays, Fridays; and at 5pm on Tuesdays and Thursdays

```{r meditation_class_size}
free_meditation %>%
  filter(Participants_Include_Instructor >= 5) %>%
  mutate(Class_Size = Participants_Include_Instructor,
         Camera = Camera_On_Participants / Participants_Include_Instructor * 100,
         Share = Sharing_Participants / Participants_Include_Instructor * 100,
         Chat = Participants_Asking_Questions_Afterwards_Commenting_In_Chat / Participants_Include_Instructor * 100) %>%
  pivot_longer(cols = c('Camera', 'Share', 'Chat')) %>%
  # plot class_size vs. measures of engagement
  ggplot(aes(x = Class_Size, y = value)) +
  geom_jitter(aes(color = name), alpha = 0.8) +
  geom_smooth(aes(color = name), method = "lm") +
  scale_y_log10() +
  scale_x_log10() +
  scale_color_viridis(discrete = TRUE) +
  labs(
    title = "Smaller Class Size is Correlated with Greater Individual Engagement",
    tag = "Figure 5",
    x = "Class Size",
    y = "Percentage of Participants",
    caption = "Each jittered dot represents one class. \n 
    Curve fits are based on linear models after log transformation.",
    color = "Engagement Type") +
  theme_gray()
```

* Figure 5 shows that smaller class sizes are correlated with higher percentages of participants taking part in some form of community engagement

* The 3 forms of community engagement considered are: turning the camera on ("Camera"), commenting in the chat window or asking questions after class ("Chat"), and sharing their experience about the meditation practice ("Share")

* This indicates that smaller group settings may encourage more active engagement by participants

* Smaller classes may lead to a higher quality of meditation practice and greater community engagement


## Testimonials

### Carey Huh

My name is Carey Huh and I worked as a neuroscientist at UCI between 2017 and 2022. My mindfulness journey began with a free virtual Mindfulness Based Stress Reduction course offered by SSIHI in May 2020, at the start of the COVID-19 pandemic. Since then, I have been a regular practitioner at SSIHI's weekday community meditations and yoga for well-being classes. The pandemic years have been a difficult, and deeply isolating, time for many of us. The community practices, albeit virtual, have provided a much needed opportunity to pause, check in, be in the present, in a shared experience with others. They certainly have been a lifesaver for my mental and physical well-being! In April 2021, I participated in the 8-week Mindful Self-Compassion course offered by SSIHI, taught by Amy Noelle. Consistent mindfulness practices and self-compassion teachings have helped to completely transform my life. I changed my career and became a data scientist working in the public health sector.

